\section{System Design}

This document outlines the architectural decisions and design patterns used in Democritus, a distributed Monte Carlo simulation engine.

\subsection{High-Level Architecture}

Democritus follows the \textbf{Master-Worker} (also known as Master-Slave) distributed computing pattern. This architecture was chosen for its simplicity, scalability, and fault-tolerance properties.

\begin{verbatim}
+-------------------+       gRPC        +-------------------+
|    SCHEDULER      |<----------------->|     WORKER 1      |
|    (Master)       |<----------------->|     WORKER 2      |
|                   |<----------------->|     WORKER N      |
+-------------------+                   +-------------------+
        |
        | HTTP /metrics
        v
+-------------------+       HTTP        +-------------------+
|   PROMETHEUS      |<------------------|    DASHBOARD      |
|   (Monitoring)    |                   |    (Next.js)      |
+-------------------+                   +-------------------+
\end{verbatim}

\subsection{Component Responsibilities}

\subsubsection{Scheduler (The Brain)}
The Scheduler is the central coordinator responsible for:
\begin{itemize}
    \item \textbf{Task Queue Management}: Maintains a queue of 10,000+ simulation tasks.
    \item \textbf{Task Distribution}: Assigns tasks to workers on-demand (pull-based model).
    \item \textbf{Result Aggregation}: Collects and stores completed simulation results.
    \item \textbf{Fault Recovery}: Implements a ``Reaper'' mechanism to reclaim timed-out tasks.
    \item \textbf{Metrics Export}: Exposes Prometheus-compatible metrics on port 2112.
\end{itemize}

\subsubsection{Worker (The Muscle)}
Workers are stateless compute nodes responsible for:
\begin{itemize}
    \item \textbf{Task Execution}: Runs the physics simulation (e.g., 3D Random Walk).
    \item \textbf{Result Submission}: Returns computed coordinates to the Scheduler.
    \item \textbf{Self-Registration}: Announces presence to the Scheduler on startup.
\end{itemize}

\subsubsection{Dashboard (The Eyes)}
A Next.js web application providing:
\begin{itemize}
    \item \textbf{Real-time Visualization}: 3D WebGL particle cloud using Three.js.
    \item \textbf{Metrics Display}: Live task completion counter and worker status.
\end{itemize}

\subsection{Design Decisions}

\subsubsection{Why gRPC over REST?}
We chose gRPC with Protocol Buffers for inter-service communication for several reasons:
\begin{enumerate}
    \item \textbf{Performance}: Binary serialization is 3-10x faster than JSON.
    \item \textbf{Strong Typing}: Protobuf schemas provide compile-time type safety.
    \item \textbf{Streaming Support}: gRPC supports bidirectional streaming (future enhancement).
    \item \textbf{Code Generation}: Automatic client/server stub generation reduces boilerplate.
    \item \textbf{Low Latency}: HTTP/2 multiplexing reduces connection overhead.
\end{enumerate}

\subsubsection{Why Pull-Based Task Distribution?}
Workers \textit{request} tasks from the Scheduler rather than having tasks \textit{pushed} to them:
\begin{itemize}
    \item \textbf{Backpressure}: Workers only request work when ready, preventing overload.
    \item \textbf{Simplicity}: No need for the Scheduler to track worker availability.
    \item \textbf{Fault Tolerance}: Dead workers simply stop requesting; no push failures.
\end{itemize}

\subsubsection{Why Stateless Workers?}
Workers maintain no persistent state between tasks:
\begin{itemize}
    \item \textbf{Horizontal Scaling}: Add/remove workers without coordination.
    \item \textbf{Failure Recovery}: Crashed workers can be replaced instantly.
    \item \textbf{Reproducibility}: All state comes from the task payload (including RNG seed).
\end{itemize}

\subsection{Data Flow}

The lifecycle of a simulation task follows this sequence:

\begin{enumerate}
    \item \textbf{Initialization}: Scheduler creates 10,000 tasks with unique seeds.
    \item \textbf{Registration}: Worker calls \texttt{RegisterWorker()} to join the grid.
    \item \textbf{Task Request}: Worker calls \texttt{GetTask()} to fetch work.
    \item \textbf{Execution}: Worker runs the simulation using the provided seed.
    \item \textbf{Submission}: Worker calls \texttt{SubmitResult()} with final coordinates.
    \item \textbf{Visualization}: Dashboard polls metrics and displays results.
\end{enumerate}

\[
\text{Worker} \xrightarrow{\texttt{GetTask()}} \text{Scheduler} \xrightarrow{\texttt{TaskPayload}} \text{Worker} \xrightarrow{\texttt{SubmitResult()}} \text{Scheduler}
\]

\subsection{Networking Architecture}

All services communicate over a Docker virtual network called \texttt{grid}:

\begin{center}
\begin{tabular}{|l|l|l|}
\hline
\textbf{Service} & \textbf{Port} & \textbf{Protocol} \\
\hline
Scheduler (gRPC) & 50051 & HTTP/2 (gRPC) \\
Scheduler (Metrics) & 2112 & HTTP/1.1 \\
Prometheus & 9090 & HTTP/1.1 \\
Dashboard & 3000 & HTTP/1.1 \\
\hline
\end{tabular}
\end{center}

Workers discover the Scheduler via Docker DNS resolution (\texttt{scheduler:50051}).

\subsection{Scalability Considerations}

\subsubsection{Horizontal Scaling}
The system scales horizontally by adding more workers:
\begin{verbatim}
docker compose up --scale worker=50
\end{verbatim}

\subsubsection{Vertical Scaling}
Individual workers can utilize multiple CPU cores via Go's goroutine scheduler.

\subsubsection{Bottlenecks}
The Scheduler is currently a single point of coordination. Future improvements could include:
\begin{itemize}
    \item Sharded task queues
    \item Multiple Scheduler replicas with consensus
    \item Distributed task storage (Redis, etcd)
\end{itemize}

\subsection{Security Considerations}

Current implementation prioritizes simplicity over security:
\begin{itemize}
    \item \textbf{No Authentication}: Workers are trusted by default.
    \item \textbf{No Encryption}: gRPC traffic is unencrypted (plaintext).
    \item \textbf{Internal Network}: All services run on a private Docker network.
\end{itemize}

For production deployment, consider:
\begin{itemize}
    \item mTLS for gRPC connections
    \item Worker authentication tokens
    \item Network policies (Kubernetes NetworkPolicy)
